---
title: 'Correlation of the altitude and number of speakers: case of Daghestan resettlement
  in XX century'
author: "L. Kazakevich, G. Moroz"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, fig.width= 12, fig.height = 9, warning = FALSE, message = FALSE)
```

```{r, message = FALSE}
library(tidyverse)
theme_set(theme_bw())
options(scipen = 999)
```


## Get dataset

A dataset provided by [Samira Verhees 2019] could be found [here](https://github.com/sverhees/master_villages). 

[A repository](https://github.com/sverhees/master_villages) provided by Samira Verhees contains a dataset with all East Caucasian villages, their coordinates and the language spoken there. We merged it with dataset provided by Yury Koryakov which contains information about number of speakers in each village according to censuses in Russia and Soviet Union (add links...).

The result dataset has the following structure:

```{r, message=FALSE}
df <- read_csv("data/merged_all_census_and_samira.csv")
df %>% 
  filter(Republic == "Daghestan") %>% 
  distinct()->
  filtered_df

filtered_df %>% 
  write_csv("data/filtered_data.csv", na = "")

str(filtered_df)
```

* `koryakov_db_id` -- id info from Y. Koryakov database
* `eng_vil_name` -- English village name from Samira's db
* `rus_vil_name` -- Russian village name from Y. Koryakov's db
* `lang_1_K` -- lanuage according to Y. Koryakov
* `lang_2_K` -- dialect info according to Y. Koryakov
* `lang_1_S` -- language according to S. Verhess
* `lang_2_S` -- dialect info according to S. Verhess
* `same_languages` -- compare  lang_1_K and lang_1_S; need to be checked by someone; after this check we will come up with a new column with language information
* `name_1885` -- village name from 1885 census (from Y. Koryakov)
* `name_1895` -- village name from 1895 census (from Y. Koryakov)
* `name_1926` -- village name from 1926 census (from Y. Koryakov)
* `name_2010` -- village name from 2010 census (from Y. Koryakov)
* `census_1885` -- number of people from 1885 census (from Y. Koryakov)
* `census_1895` -- number of people from 1895 census (from Y. Koryakov)
* `census_1926` -- number of people from 1926 census (from Y. Koryakov)
* `census_2010` -- number of people from 2010 census (from Y. Koryakov)
* `Lat` -- latitude (from S. Verhess)
* `Lon` -- longitude (from S. Verhess)
* `Republic` --  info about Republic (from S. Verhees)
* `elevation` -- elevation data (collected by G. Moroz from Google)

## Nationality by language

First we decided to look is there any patterns in census based on natianlity. We decided to remove big sities since there is no any good infromation about languages in use:

```{r}
df <- read_csv("data/filtered_data.csv")
df %>% 
  filter(lang_1_K != "Nogai") ->
  df

df %>% 
  filter(!str_detect(rus_vil_name, "Махачкала|Дербент|Буйнакск|Хасавюрт|Каспийск|Избербаш|Кизляр|Кизилюрт|Дагестанские Огни|Южно-Сухокумск")) %>%  # remove big cities
  select(koryakov_bd_id, rus_vil_name, lang_1_K, census_1885, census_1895, census_1926, census_2010) %>% 
  filter(!is.na(lang_1_K)) %>% 
  group_by(lang_1_K) %>% 
  summarise(c_1885 = sum(census_1885, na.rm = TRUE),
            c_1895 = sum(census_1895, na.rm = TRUE),
            c_1926 = sum(census_1926, na.rm = TRUE),
            c_2010 = sum(census_2010, na.rm = TRUE),
            sum = sum(c_1885, c_1895, c_1926, c_2010)) %>% 
  pivot_longer(names_to = "census", values_to = "n_people", c_1885:c_2010) %>% 
  mutate(lang_1_K = reorder(lang_1_K, sum)) %>% 
  ggplot(aes(lang_1_K, n_people, color = census, group = census))+
  geom_point()+
  geom_line()+
  coord_flip()+
  scale_y_log10()+
  labs(x = "", y = "number of people (log scale)")
```

As we can see, Russians have a peack in early censuses -- it possible that they haven't been living in some settlement at this time. It is possible to see some picks of the last census:

* amount of Kumyk, Azerbaijani, Russian, Botlikh, Tat and Hinuq speakers is significantly increased comparing to other languages
* amount of Lak a Nogai people significantly decreased
* all others are more or less the same.

Sometimes the increasing of speakers reported by census could be correlated with some patriotic movements, that e. g. resulted the increase of Botlikh (Samira Verhees personal communication).

## Differences in census
```{r}
df %>% 
  filter(!str_detect(rus_vil_name, "Махачкала|Дербент|Буйнакск|Хасавюрт|Каспийск|Избербаш|Кизляр|Кизилюрт|Дагестанские Огни|Южно-Сухокумск")) %>%  # remove big cities
  select(koryakov_bd_id, rus_vil_name, lang_1_K, census_1885, census_1895, census_1926, census_2010) %>% 
  filter(!is.na(lang_1_K)) %>% 
  group_by(lang_1_K) %>% 
  summarise(c_1895_1885 = sum(census_1895, na.rm = TRUE) - sum(census_1885, na.rm = TRUE),
            c_1926_1895 = sum(census_1926, na.rm = TRUE)- sum(census_1895, na.rm = TRUE),
            c_2010_1926 = sum(census_2010, na.rm = TRUE) - sum(census_1926, na.rm = TRUE),
            sum = sum(census_1885, census_1895, census_1926, census_2010)) %>% 
    pivot_longer(names_to = "census_diff", values_to = "n_people", c_1895_1885:c_2010_1926) %>% 
  mutate(lang_1_K = reorder(lang_1_K, sum)) %>% 
  ggplot(aes(lang_1_K, n_people)) +
  geom_point()+
  coord_flip()+
  labs(x = "", y = "difference in number of people")+
  facet_wrap(~census_diff, scales = "free_x")
```

long discription of results...

Lets have a look at the correlation between elevation and census data

```{r, fig.height=25, warning=FALSE}
df %>% 
  filter(!str_detect(rus_vil_name, "Махачкала|Дербент|Буйнакск|Хасавюрт|Каспийск|Избербаш|Кизляр|Кизилюрт|Дагестанские Огни|Южно-Сухокумск")) %>%  # remove big cities
  select(koryakov_bd_id, rus_vil_name, lang_1_K, census_1885, census_1895, census_1926, census_2010, elevation) %>% 
  filter(!is.na(lang_1_K)) %>% 
  mutate(c_1895_1885 = census_1895 - census_1885,
            c_1926_1895 = census_1926 - census_1895,
            c_2010_1926 = census_2010 - census_1926) %>% 
  select(-c(census_1885, census_1895, census_1926, census_2010)) %>% 
  pivot_longer(names_to = "census_diff", values_to = "n_diff_people", c_1895_1885:c_2010_1926)  %>% 
  ggplot(aes(n_diff_people, elevation)) +
  geom_point()+
  geom_smooth(method = "lm") +
  facet_grid(lang_1_K~census_diff)+
  theme(legend.position = "bottom")
```

If there are more points greater then 0, then amount of people increase, otherwise -- decrease. If there were any effect of elivation on number of speakers, then we would see some trend on a graph (a lot of points evenly distributed alongside blue line that should be not parallel with axis x). This could be seen only in case of Avar in differences from census 1926 to census 2010.

Conclussion: according to data there is no any clear connection between elivation and resettlement.

## map

Is there any clear areal pattern of negative or positive value for resettlement? We use only differences from census 1926 to census 2010.

```{r, cache=TRUE}
df %>% 
  filter(!str_detect(rus_vil_name, "Махачкала|Дербент|Буйнакск|Хасавюрт|Каспийск|Избербаш|Кизляр|Кизилюрт|Дагестанские Огни|Южно-Сухокумск")) %>%  # remove big cities
  select(koryakov_bd_id, rus_vil_name, lang_1_K, census_1885, census_1895, census_1926, census_2010, elevation, Lat, Lon) %>% 
  filter(!is.na(lang_1_K),
         !str_detect(lang_1_K, "тат+"),
         !is.na(Lat),
         !is.na(Lon)) %>% 
  mutate(c_2010_1926 = census_2010 - census_1926,
         positive = ifelse(c_2010_1926 > 0, "population increased", "population decreased")) ->
  df_for_plot

library(lingtypology)
map <- map.feature(languages = df_for_plot$lang_1_K,
                   latitude = df_for_plot$Lat,
                   longitude = df_for_plot$Lon,
                   features = df_for_plot$c_2010_1926,
                   facet = df_for_plot$positive)
  
library(mapview)
sync(map, no.initial.sync = FALSE)
```

It looks like that the distiributions of auls where population increased/decreased don't have any areal pattern. But it could be said, that the only strange area are sea coast (that has better life conditions) and Russian-Georgian and Russian-Azerbaijani borders. Why borders?..